\bf Interactive Visualization of Hierarchically Structured Data

### 2017-07-12 treelapse.tex
add table of problem sizes

To unify the tree-structured differential abundance and bacterial dynamics
problems, we identify the data with a collection of random variables indexed by
nodes in a prespecified tree structure. In the differential abundance problem,
each random variable lives in \(\mathbb{R}^{G}\) where \(G\) is the number of
groups being compared. Each coordinate represents the abundance for that group,
and a node exhibits differential abundance when the coordinates are drawn from
different distributions. On the other hand, in the bacterial dynamics problem,
each random variable is a time series, living in \(\mathbb{R}^{T}\).

@@

To be practically useful, the resulting visualizations must respond fluidly to
user interaction. As the data increase in scale, this fluidity can deteriorate
for two reasons. First, rendering many SVG elements in a framework like D3 is
costly\citep{johnson2008scalability}. While it is possible to use alternatives
-- HTML5's Canvas, for example -- it is often more challenging to implement
complex interactive behavior through them. Second, some of the dynamic queries
require a search over a many elements. These limitations are most pronounced in
the timebox tree display, which must search through all timepoints among all
time series whenever the brush is moved. The first, but not the second, concern
applies to treeboxes, while neither applies to DOI trees. Nonetheless, we feel
comfortable recommending timebox trees for data on the order of 500 tree tips
and 50 timepoints We note the sizes of the data sets in each case study below,
which each render fluidly, with the possible exception of the California housing
data, where there is a noticeable lag in the timebox tree and treebox displays.

@@

>>>>

To unify the tree-structured differential abundance and bacterial dynamics
problems, we identify the data with a collection of random variables indexed by
nodes in a prespecified tree structure. In the differential abundance problem,
each random variable lives in $\mathbb{R}^{G}$ where $G$ is the number of
groups being compared. Each coordinate represents the abundance for that group,
and a node exhibits differential abundance when the coordinates are drawn from
different distributions. On the other hand, in the bacterial dynamics problem,
each random variable is a time series, living in $\mathbb{R}^{T}$.

@@

To be practically useful, the resulting visualizations must respond fluidly to
user interaction. As the data increase in scale, this fluidity can deteriorate
for two reasons. First, rendering many SVG elements in a framework like D3 is
costly\citep{johnson2008scalability}. While it is possible to use alternatives
-- HTML5's Canvas, for example -- it is often more challenging to implement
complex interactive behavior through them. Second, some of the dynamic queries
require a search over a many elements. These limitations are most pronounced in
the timebox tree display, which must search through all timepoints among all
time series whenever the brush is moved. The first, but not the second, concern
applies to treeboxes, while neither applies to DOI trees. Nonetheless, we feel
comfortable recommending timebox trees for data on the order of 500 tree tips
and 50 timepoints We note the sizes of the data sets in each case study below,
which each render fluidly, with the possible exception of the California housing
data, where there is a noticeable lag in the timebox tree and treebox displays.
In problems of larger size, we recommend a preliminary filtering or aggregation
step across nodes or, if the time series is smooth, across neighboring
timepoints, in order to avoid these potential scaling difficulties.

@@

\begin{table}
\centering
\begin{tabular}{lll}
Data            & Number of timepoints & Number of nodes \\
Antibiotics     & 56                 & 386             \\
Preterm births  & 216                & 318             \\
Housing prices  & 254                & 944             \\
Bikesharing     & 24                 & 819             \\
Global patterns & 500                & 51             
\end{tabular}
\caption{Problem dimensions for each of the case studies. For problems with
  dimensions larger than that in the housing prices problem, we recommend an
  initial summarization or filtering step to prevent performance
  issues. \label{problem-scaling}}
\end{table}
